A note on the instability and degeneracy of deep learning models

### 2016-11-27 writing/draft.Rmd
sentences about applied instability and index

In other words, if the probability ratio (\ref{eqn:elpr}) is too large, then a RBM model sequence will exhibit large probability shifts for very small changes in data configurations (i.e., will exhibit instability).

>>>>

In other words, if the probability ratio (\ref{eqn:elpr}) is too large, then a RBM model sequence will exhibit large probability shifts for very small changes in data configurations (i.e., will exhibit instability). Recall the applied example of RBM models as a means to classify images. This instability result manifests itself as a one pixel change in an image (one component of the visible vector) resulting in a large shift in the probability, which in turn results in a vastly different classification of the image. Examples of this behavior have been presented in @szegedy2013intriguing for trained deep learning models, in which a one pixel change in a test image results in a wildly different classification.

### 2016-11-27 writing/note.Rmd
sentences about applied instability and index

For $\boldsymbol X = X_1, \dots, X_N$ discrete random variables with sample space $\mathcal{X}^N$, $|\mathcal{X}| < \infty$, consider an exponential family model for $\boldsymbol X$ with probability mass function of the form
$$
p_{\boldsymbol \lambda}(\boldsymbol x) = \exp\left[\boldsymbol\eta^T(\boldsymbol \lambda) \boldsymbol g(\boldsymbol x) - \psi(\boldsymbol \lambda)\right], \qquad \boldsymbol x \in \mathcal{X}^N,
$$ 
for fixed positive integers $k$ and $L$ denoting the dimensions of the parameter and natural parameters spaces respectively, $\boldsymbol \lambda \in \Lambda \subset \mathbb{R}^{k}$, $\boldsymbol \eta : \mathbb{R}^k \mapsto \mathbb{R}^L$, $\boldsymbol g : \mathcal{X}^N \mapsto \mathbb{R}^L$ a vector of sufficient statistics, 
$$
\psi(\boldsymbol \lambda) = \log \sum\limits_{\boldsymbol x \in \mathcal{X}^N}\exp\left[\boldsymbol \eta^T(\boldsymbol \lambda) \boldsymbol g(\boldsymbol x) \right], \qquad \boldsymbol \lambda \in \Lambda,
$$
the normalizing function, and $\Lambda = \{\boldsymbol \lambda \in \mathbb{R}^k : \psi(\boldsymbol \lambda) < \infty, k \le q(N) \}$ is the parameter space.

Defining $\boldsymbol \theta_N=\lambda_N$ to a sequence of elements of $\Lambda \subset \mathbb{R}^k$ and noting that $p_{\boldsymbol \lambda}(\boldsymbol x) > 0$ for all $\boldsymbol x \in \mathcal{X}^N$, these discrete exponential family models are special cases of the FSFS models. Such exponential models arise with spatial data on a lattice [@besag1974spatial], network data [@wasserman1994social; @handcock2003assessing], and even binomial sampling with $N$ iid Bernoulli random variables. For these exponential models, the dimension of the parameter $\boldsymbol \theta_N$ is the same for each $N$ (i.e., $\boldsymbol \theta_N$ lies in a parameter space of fixed Euclidean dimension $k$).  This need not be true for other types of FSFS models considered in Sections \ref{restricted-boltzmann-machines}-\ref{deep-learning}. @schweinberger2011instability considered instability in such exponential models (e.g., for random graphs) for sequences of fixed parameters $\theta_N=\lambda\in\mathbb{R}^k$, $N \geq 1$, of non-varying dimension $k$.

>>>>

For $\boldsymbol X = X_1, \dots, X_N$ discrete random variables with sample space $\mathcal{X}^N$, $|\mathcal{X}| < \infty$, consider an exponential family model for $\boldsymbol X$ with probability mass function of the form
$$
p_{N, \boldsymbol \lambda}(\boldsymbol x) = \exp\left[\boldsymbol\eta^T(\boldsymbol \lambda) \boldsymbol g_N(\boldsymbol x) - \psi(\boldsymbol \lambda)\right], \qquad \boldsymbol x \in \mathcal{X}^N,
$$ 
for fixed positive integers $k$ and $L$ denoting the dimensions of the parameter and natural parameters spaces respectively, $\boldsymbol \lambda \in \Lambda \subset \mathbb{R}^{k}$, $\boldsymbol \eta : \mathbb{R}^k \mapsto \mathbb{R}^L$, $\boldsymbol g_N : \mathcal{X}^N \mapsto \mathbb{R}^L$ a vector of sufficient statistics, 
$$
\psi(\boldsymbol \lambda) = \log \sum\limits_{\boldsymbol x \in \mathcal{X}^N}\exp\left[\boldsymbol \eta^T(\boldsymbol \lambda) \boldsymbol g_N(\boldsymbol x) \right], \qquad \boldsymbol \lambda \in \Lambda,
$$
the normalizing function, and $\Lambda = \{\boldsymbol \lambda \in \mathbb{R}^k : \psi(\boldsymbol \lambda) < \infty, k \le q(N) \}$ is the parameter space.

Defining $\boldsymbol \theta_N=\lambda$ to a sequence of elements of $\Lambda \subset \mathbb{R}^k$ and noting that $p_{N, \boldsymbol \lambda}(\boldsymbol x) > 0$ for all $\boldsymbol x \in \mathcal{X}^N$, these discrete exponential family models are special cases of the FSFS models. Such exponential models arise with spatial data on a lattice [@besag1974spatial], network data [@wasserman1994social; @handcock2003assessing], and even binomial sampling with $N$ iid Bernoulli random variables. For these exponential models, the dimension of the parameter $\boldsymbol \theta_N$ is the same for each $N$ (i.e., $\boldsymbol \theta_N$ lies in a parameter space of fixed Euclidean dimension $k$).  This need not be true for other types of FSFS models considered in Sections \ref{restricted-boltzmann-machines}-\ref{deep-learning}. @schweinberger2011instability considered instability in such exponential models (e.g., for random graphs) for sequences of fixed parameters $\theta_N=\lambda\in\mathbb{R}^k$, $N \geq 1$, of non-varying dimension $k$.
